\bf Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2025-06-10 manuscript/manuscript.tex
edits and comments

Developing a fully nonparametric solution for the KS
tests in the presence of serial dependence presents a significant
challenge. The null distribution's characteristics are intricately linked to
the structure of the serial dependence, which can vary widely in practical
situations. Employing a parametric method would necessitate defining a
specific model for this dependence, despite the primary focus being on
evaluating the marginal distribution of a stationary series. To date, a bias
correction method for block bootstrap, akin to the NPB
approach of \citet{babu2004goodness} for independent data,
has not been established. This study seeks
to bridge this gap by introducing a bias correction technique for the
nonparametric block bootstrap (NPBB), tailored for use in scenarios where the
hypothesized distribution has unspecified parameters and the data
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

@@

Assume that the time series $\{X_i\}$ is stationary and
strongly mixing. Using arguments similar to those
in \citet{kunsch1989jackknife}, under certain conditions, the
following results can be shown to
justify the validity of the NPBB procedure of the KS test.
As $1/l+l/n\rightarrow 0$,
\[
  T_n \overset{d}{\rightarrow}
  \sup_x \big\vert G_{\infty}(x)+f(x;\theta_0)U \big\vert
\]
for some appropriately defined Gaussian process $G_{\infty}(\cdot)$
and Gaussian random variable  $U$.
Further,  as $1/l+l/n\rightarrow 0$,
\[
  \big\vert \mathbb{P}(T_n^{(b)}\leq x \mid \{X_i: i = 1, \ldots, n\}) -
  \mathbb{P}(T_n\leq x) \big\vert
  \overset{a.s.}{\rightarrow} 0
\]
for any fixed $x$, where $T_n^{(b)}$ is the statistic computed
based on the bootstrap sample $\{X_i^{(b)}: i = 1, \ldots, n\}$.
Following the arguments in the proof of the Glivenko--Cantelli
Theorem, the bootstrap consistency result could be further
strengthened to allow uniform convergence over $x$. That is,
as $1/l+l/n\rightarrow 0$,
\[
  \sup_x \big \vert
  \mathbb{P}(T_n^{(b)}\leq x \mid \{X_i: i = 1, \ldots, n\}) - \mathbb{P}(T_n\leq x) \big\vert
  \overset{a.s.}{\rightarrow} 0.
\]

@@

Using the \textsl{tseries} package \citep{tseries},
we gathered daily closing prices for the S\&P 500 index from January 1st, 2020
to December 31st, 2023. Daily returns were obtained by taking the difference in
the logarithm of the closing prices.
It is known as stylized facts that daily return series often exhibit nonlinear
dependence and unconditional heavy tails \citep[e.g.,][]{ryden1998stylized,
  cont2001empirical}. The lag-1 Kendall's $\tau$ of the series is
about $-0.0298$, indicating a fairly weak serial linear correlation, but
certainly not meaning absence of serial dependence.
Using our method, we tested that the daily returns are normally
distributed and Student's $t$ distributed
with degrees of freedom $\nu \in \{30, 20, 10, 5, 4, 3, 2, 1\}$
with the proposed NPBB method.
For comparison, we tested the same hypotheses using 1) the
SPB method of \citet{zeimbekakis2022misuses},
which accounts for both serial dependence and unspecified parameters;
2) NPB bias correction method of
\citet{babu2004goodness}; and 3) the parametric bootstrap (PB) method shown
by \citet{zeimbekakis2022misuses}. The latter of the two methods discard the
serial dependence. In all methods, we used $B = 10,000$ bootstrap replicates.
The p-values for these tests are summarized in
Table~\ref{table:SP5004}.

@@

While differences in p-values among different methods are expected, our analysis
reveals some instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2022misuses}.
For example, our NPBB method fails to reject the
hypothesis that the series follows a Student's $t$ distribution with degrees of
freedom $v = 5$ with a p-value of 0.061. In contrast,
the SPB method yields a p-value of 0.0192,
the NPB method yields a p-value of 0.0222, and the PB yields a p-value of
0.0141. Although such differences are not guaranteed, this
instance highlights a disagreement in conclusions among methods for the S\&P 500
index. Because NPB and PB both discard serial dependence, it is expected that
their conclusions could be different from our method. It is worth noting that
the SPB result is also not reliable here because the working dependence model is
in the class of ARMA dependence as specified in \citet{zeimbekakis2022misuses},
which does not cover nonlinear dependence usually captured by autoregressive
conditional heteroskedasticity \citep{engle1995arch}. The proposed method has
no need to specify the serial dependence and, hence, produces results under
weakest assumptions.

@@

The KS test remains a crucial tool for researchers across diverse fields,
enabling critical assessments of population distributions. We introduced a
NPBB variant of the KS test tailored for data with
serial dependencies, which eliminates the need for explicit modeling serial
dependence in addition to parameter specification within the hypothesized
distribution. Our simulations have demonstrated that, given
sufficiently large samples of time series data and moderate serial dependencies,
the test maintains its size under the null hypothesis and have considerable
power in detecting deviations from the hypothesized marginal distribution. The
method's performance improves as the strength of temporal dependence decreases
or the sample size increases. For moderate serial dependence levels in practice,
the test can be recommended for sample size~$n = 400$ or higher.
Overall, the test provides an analog of the bias-corrected NPB KS test of
\citet{babu2004goodness} in the context of stationary time
series with unspecified serial dependence. Compared to the SPB method
of \citet{zeimbekakis2022misuses}, which requires a working dependence model, it
is reliable with no need to worry about whether the working dependence
model approximates the true dependence model.

@@

\section{Block Size Investigation}\label{sec:appendix}

@@

Still, even for settings where the block sizes are usually less than $\sqrt{n}$ -
such as for $n = 800$, $\tau = 0.5$, and $X_i \sim N(8, 8)$ - the 
block sizes are almost always larger than $l = \lceil n^{1/3} \rceil$, a setting
which was considered optimal by \citep{buhlmann1999block}. This
is demonstrated in Figure~\ref{fig:block_dist} with a histogram
of the selected block sizes for this setting.

@@

Despite this, the
algorithm with \citet{politis2004automatic}'s setting holds its size somewhat 
well for moderately 
positively dependent
series as demonstrated in 
Figure~\ref{fig:alt_qq}. Both 
\citet{politis2004automatic}'s setting and the cubic root setting result
in the method not holding its size for $|\tau| > 0.5$. 
Figure~\ref{fig:alt_rr} also shows that the two settings are comparably
powerful for $|\tau| <= 0.5$, with \citet{politis2004automatic}'s
setting perhaps even having higher rejection rates for smaller negatively
dependent sample sizes. 
However, we still
recommend the original cubic root setting over that introduced by 
\citet{politis2004automatic} because the latter fails to hold its size
for negative $\tau$, particularly for smaller sample sizes.

>>>>

Developing a fully nonparametric solution for the KS
tests in the presence of serial dependence presents a significant
challenge. The null distribution's characteristics are intricately linked to
the structure of the serial dependence, which can vary widely in practical
situations. Employing a semiparametric method
\citep{zeimbakakis2024misuses} would necessitate defining a
specific model for this dependence, despite the primary focus being on
evaluating the marginal distribution of a stationary series. To date, a bias
correction method for block bootstrap, akin to the NPB
approach of \citet{babu2004goodness} for independent data,
has not been established. This study seeks
to bridge this gap by introducing a bias correction technique for the
nonparametric block bootstrap (NPBB), tailored for use in scenarios where the
hypothesized distribution has unspecified parameters and the data
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

@@

Assume that the time series $\{X_i\}$ is stationary and
strongly mixing. Using arguments similar to those
in \citet{kunsch1989jackknife}, under certain conditions, the
following results can be shown to
justify the validity of the NPBB procedure of the KS test.
As $1/l+l/n\rightarrow 0$,
\[
  T_n \overset{d}{\rightarrow}
  \sup_x \big\vert G_{\infty}(x)+f(x;\theta_0)U \big\vert
\]
\jy{What is $f$? Be careful with the dimension of $U$ and $f$.}
for some appropriately defined Gaussian process $G_{\infty}(\cdot)$
and Gaussian random variable  $U$.
Further,  as $1/l+l/n\rightarrow 0$,
\[
  \big\vert \mathbb{P}(T_n^{(b)}\leq x \mid \{X_i: i = 1, \ldots, n\}) -
  \mathbb{P}(T_n\leq x) \big\vert
  \overset{a.s.}{\rightarrow} 0
\]
for any fixed $x$, where $T_n^{(b)}$ is the statistic computed
based on the bootstrap sample $\{X_i^{(b)}: i = 1, \ldots, n\}$.
Following the arguments in the proof of the Glivenko--Cantelli
Theorem, the bootstrap consistency result could be further
strengthened to allow uniform convergence over $x$. That is,
as $1/l+l/n\rightarrow 0$,
\[
  \sup_x \big \vert
  \mathbb{P}(T_n^{(b)}\leq x \mid \{X_i: i = 1, \ldots, n\}) - \mathbb{P}(T_n\leq x) \big\vert
  \overset{a.s.}{\rightarrow} 0.
\]

@@

Using the \textsl{tseries} package \citep{tseries},
we gathered daily closing prices for the S\&P 500 index from January 1st, 2020
to December 31st, 2023. Daily returns were obtained by taking the difference in
the logarithm of the closing prices.
It is known as stylized facts that daily return series often exhibit nonlinear
dependence and unconditional heavy tails \citep[e.g.,][]{ryden1998stylized,
  cont2001empirical}. The lag-1 Kendall's $\tau$ of the series is
about $-0.0298$, indicating a fairly weak serial linear correlation, but
certainly not meaning absence of serial dependence.
Using our method, we tested that the daily returns are normally
distributed and Student's $t$ distributed
with degrees of freedom $\nu \in \{30, 20, 10, 5, 4, 3, 2, 1\}$
with the proposed NPBB method.
For comparison, we tested the same hypotheses using 1) the
SPB method of \citet{zeimbekakis2024misuses},
which accounts for both serial dependence and unspecified parameters;
2) NPB bias correction method of
\citet{babu2004goodness}; and 3) the parametric bootstrap (PB) method shown
by \citet{zeimbekakis2024misuses}. The latter of the two methods discard the
serial dependence. In all methods, we used $B = 10,000$ bootstrap replicates.
The p-values for these tests are summarized in
Table~\ref{table:SP5004}.

@@

While differences in p-values among different methods are expected, our analysis
reveals some instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2024misuses}.
For example, our NPBB method fails to reject the
hypothesis that the series follows a Student's $t$ distribution with degrees of
freedom $v = 5$ with a p-value of 0.061. In contrast,
the SPB method yields a p-value of 0.0192,
the NPB method yields a p-value of 0.0222, and the PB yields a p-value of
0.0141. Although such differences are not guaranteed, this
instance highlights a disagreement in conclusions among methods for the S\&P 500
index. Because NPB and PB both discard serial dependence, it is expected that
their conclusions could be different from our method. It is worth noting that
the SPB result is also not reliable here because the working dependence model is
in the class of ARMA dependence as specified in \citet{zeimbekakis2024misuses},
which does not cover nonlinear dependence usually captured by autoregressive
conditional heteroskedasticity \citep{engle1995arch}. The proposed method has
no need to specify the serial dependence and, hence, produces results under
weakest assumptions.

@@

The KS test remains a crucial tool for researchers across diverse fields,
enabling critical assessments of population distributions. We introduced a
NPBB variant of the KS test tailored for data with
serial dependencies, which eliminates the need for explicit modeling serial
dependence in addition to parameter specification within the hypothesized
distribution. Our simulations have demonstrated that, given
sufficiently large samples of time series data and moderate serial dependencies,
the test maintains its size under the null hypothesis and have considerable
power in detecting deviations from the hypothesized marginal distribution. The
method's performance improves as the strength of temporal dependence decreases
or the sample size increases. For moderate serial dependence levels in practice,
the test can be recommended for sample size~$n = 400$ or higher.
Overall, the test provides an analog of the bias-corrected NPB KS test of
\citet{babu2004goodness} in the context of stationary time
series with unspecified serial dependence. Compared to the SPB method
of \citet{zeimbekakis2024misuses}, which requires a working dependence model, it
is reliable with no need to worry about whether the working dependence
model approximates the true dependence model.

@@

\section{Block Size Investigation}\label{sec:blocksize}

@@

\jy{Watch proper usage of dashes.}
Still, even for settings where the block sizes are usually less than $\sqrt{n}$ -
such as for $n = 800$, $\tau = 0.5$, and $X_i \sim N(8, 8)$ - the 
block sizes are almost always larger than $l = \lceil n^{1/3} \rceil$, a setting
which was considered optimal by \citep{buhlmann1999block}. This
is demonstrated in Figure~\ref{fig:block_dist} with a histogram
of the selected block sizes for this setting.
\jy{What's the logical connection of this paragraph with the
  surrounding paragraphs?}

@@

Despite this, the
algorithm with \citet{politis2004automatic}'s setting holds its size somewhat 
well for moderately 
positively dependent
series as demonstrated in 
Figure~\ref{fig:alt_qq}. Both 
\citet{politis2004automatic}'s setting and the cubic root setting result
in the method not holding its size for $|\tau| > 0.5$. 
Figure~\ref{fig:alt_rr} also shows that the two settings are comparably
powerful for $|\tau| <= 0.5$, with \citet{politis2004automatic}'s
setting perhaps even having higher rejection rates for smaller negatively
dependent sample sizes. However, we still
\jy{too strong. soften it.}
recommend the original cubic root setting over that introduced by 
\citet{politis2004automatic} because the latter fails to hold its size
for negative $\tau$, particularly for smaller sample sizes.

### 2025-06-10 manuscript/reply.tex
edits and comments

The revised manuscript incorporates the following notable changes:
\begin{enumerate}
\item An appendix has been added documenting our exploration of
  automatic block size selection for circular block bootstrap,
  including observed limitations in practice.
\item  A paragraph has been added to the Discussion section to
  discuss the impact of block size on test performance and to
  contextualize our findings in light of existing literature on
  optimal block length
\item A stronger and more detailed justification of the bias
  correction term is now provided in a newly added Section~2.4.
\end{enumerate}

@@

\begin{comment}
1. I also noticed that on Page 3 it says “Employing a nonparametric method would necessitate defining a specific model for this dependence”, but I think here “nonparametric” should be “parametric”.
\end{comment}

Fixed.

@@

\begin{itemize}
\item We have added a discussion to the conclusion section (Section~5) 
that explicitly addresses the role of block size and its effect on the 
performance of the NPBB test. This includes a review of relevant literature, 
as suggested by the reviewer, including works by 
\citet{hall1995blocking}, \citet{lahiri1999theoretical}, 
\citet{buhlmann2002bootstraps}, \citet{politis2004automatic}, and 
\citet{lahiri2013resampling}.

\item Motivated by these references, we implemented the automatic block 
length selection method proposed by \citet{politis2004automatic} for 
circular block bootstrap and reported our findings in a new appendix. 
While this method is attractive in principle, our empirical results show 
that it often selects overly large block sizes---especially under strong 
dependence---resulting in limited variability among bootstrap replicates 
and diminished test performance. These observations echo the theoretical 
concerns raised in the literature and underscore the practical limitations 
of fully automatic block length selection in the context of 
goodness-of-fit testing.
\end{itemize}

@@

\begin{comment}
2. The work presented here is useful, however I have believe the explanations
for the correction term may need modification. Here the entire justification is
based on the null case which does not cover the case when the null is not true.
Justification for the correction term in \citet{babu2004goodness} is not based 
on just the
null case. The same seems to be true here, and in that case, Section 2 needs
rewriting.

@@

\begin{align*}
Y_n(x, \hat \theta_n) &= \sqrt{n}[F_n(x) - F(x; \hat \theta_n)],\\
Y_n^{(b)}(x, \hat \theta_n^{(b)}) &= \sqrt{n}[F_n^{(b)}(x) - F(x; \hat \theta_n^{(b)})],
\end{align*}

where the second quantity is based on the $b$th bootstrap sample. In the iid 
case, \citet{babu2004goodness} introduced a correction term for the nonparametric bootstrap
$$B_n(x) = \sqrt{n}[F_n(x) - F(x; \hat \theta_n)],$$
so that
\begin{align*}
Y_{n,corr}^{(b)}(x) &= Y_n^{(b)}(x; \hat\theta_n^{(b)}) - B_n(x)\\
                    &= \sqrt{n}[F_n^{(b)}(x) - F_n(x)] -
                    \sqrt{n}[F(x; \hat \theta_n^{(b)}) -  F(x; \hat\theta_n)].
\end{align*}
It was proved in \citet{babu2004goodness} that the distribution of the KS
statistic is approximated by that of the corrected bootstrap under the null as
well as for contiguous alternatives.

@@

Thank you for drawing this to our attention. As is standard in hypothesis 
testing, our development of the bootstrap procedure is focused on accurately 
approximating the distribution of the test statistic under the null hypothesis. 
This is consistent with the goal of using the bootstrap to construct valid
 p-values and rejection regions. While the same bias correction is applied 
regardless of whether the null is true, its purpose is to replicate the 
behavior of the statistic under the null, which is the basis for valid inference.

>>>>

The revised manuscript incorporates the following notable changes:
\begin{enumerate}
\item An appendix has been added documenting our exploration of
  automatic block size selection for circular block bootstrap,
  including observed limitations in practice.
\item  A paragraph has been added to the Discussion section to
  discuss the impact of block size on test performance and to
  contextualize our findings in light of existing literature on
  optimal block size.
\item A stronger and more detailed justification of the bias
  correction term is provided in a newly added Section~2.4.
\end{enumerate}

@@

\begin{comment}
I also noticed that on Page 3 it says “Employing a nonparametric method would necessitate defining a specific model for this dependence”, but I think here “nonparametric” should be “parametric”.
\end{comment}

We intended to refer to the semiparametric bootstrap method as
discussed in \citep{zeimbekakis2024misuses}. The sentence has been
revised to clarify this point and to correctly distinguish it from a
fully nonparametric approach.

@@

\begin{itemize}
\item We have added a discussion to the \emph{Discussion} section
  (paragraph~2, Section~5) to explicitly discuss
  the role of block size and its effect on the 
  performance of the NPBB test, citing the literature suggested by the
  reviewer.

\item Motivated by these references, we implemented the automatic block 
length selection method proposed by \citet{politis2004automatic} for 
circular block bootstrap and reported our findings in a newly added
section in the Appendix. 
While this method is attractive in principle, our empirical results show 
that it often selects overly large block sizes---especially under strong 
dependence---resulting in limited variability among bootstrap replicates 
and diminished test performance. These observations echo the theoretical 
concerns raised in the literature and underscore the practical limitations 
of fully automatic block length selection in the context of 
goodness-of-fit testing.
\end{itemize}

@@

\begin{comment}
2. The work presented here is useful, however I have believe the explanations
for the correction term may need modification. Here the entire justification is
based on the null case which does not cover the case when the null is not true.
Justification for the correction term in \citet{babu2004goodness} is not based 
on just the
null case. The same seems to be true here, and in that case, Section~2 needs
rewriting.

@@

\begin{align*}
Y_n(x, \hat \theta_n) &= \sqrt{n}[F_n(x) - F(x; \hat \theta_n)],\\
Y_n^{(b)}(x, \hat \theta_n^{(b)}) &= \sqrt{n}[F_n^{(b)}(x) - F(x; \hat \theta_n^{(b)})],
\end{align*}
where the second quantity is based on the $b$th bootstrap sample. In the iid 
case, \citet{babu2004goodness} introduced a correction term for the nonparametric bootstrap
$$B_n(x) = \sqrt{n}[F_n(x) - F(x; \hat \theta_n)],$$
so that
\begin{align*}
Y_{n,corr}^{(b)}(x) &= Y_n^{(b)}(x; \hat\theta_n^{(b)}) - B_n(x)\\
                    &= \sqrt{n}[F_n^{(b)}(x) - F_n(x)] -
                    \sqrt{n}[F(x; \hat \theta_n^{(b)}) -  F(x; \hat\theta_n)].
\end{align*}
It was proved in \citet{babu2004goodness} that the distribution of the KS
statistic is approximated by that of the corrected bootstrap under the null as
well as for contiguous alternatives.

@@

Thank you for the detailed demonstration. As is standard in hypothesis 
testing, our development of the bootstrap procedure is focused on accurately 
approximating the distribution of the test statistic under the null hypothesis. 
This is consistent with the goal of using the bootstrap to construct valid
p-values and rejection regions. While the same bias correction is applied 
regardless of whether the null is true, its purpose is to replicate the 
behavior of the statistic under the null, which is the basis for valid inference.
